---
title: "University Student Survey — Executive Dashboard"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    theme: flatly
---

```{=html}
<style>
body { font-family: -apple-system, "Segoe UI", Helvetica, Arial, sans-serif; }
.navbar { font-weight: 600; letter-spacing: 0.02em; }
.chart-title {
  font-weight: 700 !important;
  font-size: 15px !important;
  color: #14202E !important;
  border-bottom: 2px solid #E4E8EC;
}
.value-box .value { font-weight: 700 !important; font-size: 30px !important; }
.value-box .caption { font-weight: 500 !important; opacity: 0.93; }
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

required_pkgs <- c("flexdashboard", "tidyverse", "readxl", "janitor",
                    "plotly", "DT", "scales")
for (pkg in required_pkgs) {
  if (!requireNamespace(pkg, quietly = TRUE)) install.packages(pkg)
}

library(flexdashboard)
library(tidyverse)
library(readxl)
library(janitor)
library(plotly)
library(DT)
library(scales)

set.seed(2026)

# ================================================================ PALETTE ===
INK   <- "#14202E"
MUTED <- "#7A8794"
LOW   <- "#B3202C"
MID   <- "#E8B33C"
HIGH  <- "#12715A"

pillar_palette <- c(
  "Facilities & Campus" = "#2C5F8A",
  "Academic Quality"    = "#12715A",
  "Digital Readiness"   = "#7B5C9E",
  "Financial Security"  = "#C75B3F",
  "Health & Wellbeing"  = "#47808C",
  "Support Services"    = "#D9A03C",
  "Career Readiness"    = "#8C4A83",
  "General Experience"  = "#7A8794"
)

band_colors_5 <- c("Critical" = "#8C1720", "At risk" = "#C75B3F",
                    "Stable" = "#D9B23C", "Thriving" = "#5FA88E",
                    "Excelling" = "#12715A")

# ============================================================================
# LOAD & CLEAN
# ============================================================================

dat <- read_excel("Book1.xlsx") %>% clean_names()

dat <- dat %>%
  select(-matches("name|respondent_id|timestamp|email|phone|enrollment|roll_no|student_id",
                  ignore.case = TRUE))

dat <- dat[rowMeans(is.na(dat)) <= 0.6, ]
dat <- dat[!duplicated(dat), ]

sector_col <- grep("sector|institution_type", names(dat), value = TRUE)[1]
gender_col <- grep("gender|sex", names(dat), value = TRUE)[1]
univ_col   <- grep("university_reported|university|institution", names(dat), value = TRUE)[1]
field_col  <- grep("field_group|discipline|faculty|program|field", names(dat), value = TRUE)[1]
age_col    <- grep("^q.*age|^age", names(dat), value = TRUE)[1]

if (!is.na(sector_col)) {
  s <- tolower(trimws(as.character(dat[[sector_col]])))
  s[grepl("govt|government|public|federal", s)] <- "Public"
  s[grepl("pvt|private", s)] <- "Private"
  s[!s %in% c("Public", "Private")] <- NA
  dat$sector <- s
} else {
  dat$sector <- NA
}

if (!is.na(gender_col)) {
  g <- tolower(trimws(as.character(dat[[gender_col]])))
  g[g %in% c("m", "male")] <- "Male"
  g[g %in% c("f", "female")] <- "Female"
  g[!g %in% c("Male", "Female")] <- NA
  dat$gender <- g
} else {
  dat$gender <- NA
}

if (!is.na(univ_col)) {
  dat$university <- str_to_title(str_squish(as.character(dat[[univ_col]])))
  dat$university[dat$university %in% c("", "Na", "N/A")] <- NA
} else {
  dat$university <- NA
}

if (!is.na(field_col)) {
  dat$field <- str_to_title(str_squish(as.character(dat[[field_col]])))
  dat$field[dat$field %in% c("", "Na", "N/A")] <- NA
} else {
  dat$field <- NA
}

if (!is.na(age_col)) {
  dat$age <- suppressWarnings(as.numeric(dat[[age_col]]))
  dat$age[dat$age < 15 | dat$age > 60] <- NA
} else {
  dat$age <- NA
}

# ============================================================================
# STRICT LIKERT DETECTION
# ============================================================================

lk <- c("strongly disagree" = 1, "disagree" = 2, "neutral" = 3,
        "neither agree nor disagree" = 3, "agree" = 4, "strongly agree" = 5,
        "very dissatisfied" = 1, "dissatisfied" = 2, "satisfied" = 4,
        "very satisfied" = 5, "very poor" = 1, "poor" = 2, "average" = 3,
        "good" = 4, "very good" = 5, "excellent" = 5, "fair" = 3,
        "never" = 1, "rarely" = 2, "sometimes" = 3, "often" = 4, "always" = 5,
        "not at all" = 1, "slightly" = 2, "moderately" = 3, "very" = 4,
        "extremely" = 5)

scale_words <- names(lk)[!grepl("^[0-9]+$", names(lk))]

BLOCK_PATTERN <- paste0(
  "sibling|semester|^age|_age$|year_of|number|count|cgpa|gpa|marks|",
  "income|amount|hours|province|district|city|university|institution|",
  "gender|sex|field|program|degree|sector|status|category|type"
)

norm_vals <- function(x) {
  v <- tolower(trimws(as.character(x)))
  v[v %in% c("", "na", "n/a", "-", "none", "nil", "null")] <- NA
  v
}

item_cols <- character(0)

for (cl in names(dat)) {
  if (cl %in% c("sector", "gender", "age", "university", "field")) next
  if (grepl(BLOCK_PATTERN, cl)) next
  if (grepl("_num$", cl)) next

  x <- dat[[cl]]
  if (!is.character(x)) next

  v <- norm_vals(x)
  nn <- sum(!is.na(v))
  if (nn < 50) next

  uniq <- unique(v[!is.na(v)])
  if (length(uniq) > 12) next
  if (!any(uniq %in% scale_words)) next

  mapped <- unname(lk[v])
  coverage <- sum(!is.na(mapped)) / nn
  if (coverage < 0.80) next

  newcol <- paste0(cl, "_num")
  dat[[newcol]] <- as.numeric(mapped)
  item_cols <- c(item_cols, newcol)
}

# ============================================================================
# PILLARS BY THEME
# ============================================================================

pillar_defs <- list(
  "Facilities & Campus" = "librar|washroom|canteen|cafeteria|sport|recreation|transport|classroom|lab|wifi|wi_fi|internet|multimedia|first_aid|hostel|building|parking|water|clean|hygien|campus|infrastructure|facilit",
  "Academic Quality"    = "syllabus|curricul|exam|grading|assess|teacher|faculty|workload|theory|practical|content|lecture|teaching|course|research|academic|impartial|feedback|knowledge",
  "Digital Readiness"   = "digital|online|computer|device|software|technolog|lms|e_learn|laptop|smartphone",
  "Financial Security"  = "financ|fee|expense|cost|money|scholarship|afford|stipend|tuition",
  "Health & Wellbeing"  = "health|stress|sleep|wellbeing|well_being|mental|anxiet|diet|exercise|illness",
  "Support Services"    = "support|guidance|advisor|mentor|counsel|help|administration|complaint|service",
  "Career Readiness"    = "career|internship|employ|job|prepare|skill|placement|entrepreneur|future"
)

assign_pillar <- function(col) {
  nm <- tolower(col)
  for (p in names(pillar_defs)) {
    if (grepl(pillar_defs[[p]], nm)) return(p)
  }
  "General Experience"
}

item_map <- tibble(var = item_cols) %>%
  mutate(Pillar = map_chr(var, assign_pillar))

pillar_counts <- item_map %>% count(Pillar, name = "n_items")
valid_pillars <- pillar_counts$Pillar[pillar_counts$n_items >= 2]
item_map <- item_map %>% filter(Pillar %in% valid_pillars)

row_index <- function(d, cols) {
  m <- suppressWarnings(apply(as.matrix(d[, cols, drop = FALSE]), 2, as.numeric))
  if (is.null(dim(m))) m <- matrix(m, ncol = 1)
  o <- rowMeans(m, na.rm = TRUE) * 20
  o[rowSums(!is.na(m)) == 0] <- NA_real_
  pmin(100, pmax(0, o))
}

pillar_vars <- character(0)
PILLAR_LABEL <- character(0)

for (p in valid_pillars) {
  cols_p <- item_map$var[item_map$Pillar == p]
  vname <- paste0("pil_", make.names(p))
  dat[[vname]] <- row_index(dat, cols_p)
  if (sum(!is.na(dat[[vname]])) >= 50) {
    pillar_vars <- c(pillar_vars, vname)
    PILLAR_LABEL[vname] <- p
  }
}

if (length(item_map$var) > 0) {
  dat$idx_overall <- row_index(dat, item_map$var)
} else {
  dat$idx_overall <- NA_real_
}

if (length(pillar_vars) > 0) {
  pillar_vars <- pillar_vars[order(-sapply(pillar_vars,
                                            function(v) mean(dat[[v]], na.rm = TRUE)))]
}

PILLARS <- pillar_vars
LAB <- PILLAR_LABEL

rate_word <- function(m) {
  case_when(m >= 80 ~ "Very good", m >= 60 ~ "Good",
            m >= 40 ~ "Needs work", TRUE ~ "Poor")
}

# ---- Headline numbers ------------------------------------------------------
total_students <- nrow(dat)
avg_overall <- if (sum(!is.na(dat$idx_overall)) > 0) {
  round(mean(dat$idx_overall, na.rm = TRUE), 1)
} else NA

n_pillars <- length(PILLARS)
n_items_total <- nrow(item_map)

uni_counts <- dat %>% filter(!is.na(university)) %>% count(university)
n_institutions <- sum(uni_counts$n >= 5)

# ---- Pillar summary with spread -------------------------------------------
pillar_stats <- NULL
if (length(PILLARS) > 0) {
  pillar_stats <- map_dfr(PILLARS, function(v) {
    x <- dat[[v]][!is.na(dat[[v]])]
    tibble(Pillar = LAB[[v]],
           Items = sum(item_map$Pillar == LAB[[v]]),
           Students = length(x),
           Index = round(mean(x), 1),
           Q1 = round(quantile(x, 0.25), 1),
           Median = round(median(x), 1),
           Q3 = round(quantile(x, 0.75), 1))
  }) %>%
    mutate(`Gap to overall` = round(Index - avg_overall, 1),
           Rating = rate_word(Index)) %>%
    arrange(desc(Index))
}

pillar_tbl <- NULL
if (!is.null(pillar_stats)) {
  pillar_tbl <- pillar_stats %>%
    select(Pillar, Items, Students, Index, `Gap to overall`, Rating)
}

best_pillar  <- if (!is.null(pillar_stats)) pillar_stats$Pillar[1] else "N/A"
best_score   <- if (!is.null(pillar_stats)) pillar_stats$Index[1] else NA
worst_pillar <- if (!is.null(pillar_stats)) pillar_stats$Pillar[nrow(pillar_stats)] else "N/A"
worst_score  <- if (!is.null(pillar_stats)) pillar_stats$Index[nrow(pillar_stats)] else NA

# ---- Risk segmentation -----------------------------------------------------
risk_tbl <- NULL
pct_at_risk <- NA
if (sum(!is.na(dat$idx_overall)) >= 50) {
  qs <- unique(quantile(dat$idx_overall, probs = c(.2, .4, .6, .8),
                         na.rm = TRUE, type = 7))
  if (length(qs) == 4) {
    dat$risk_band <- cut(dat$idx_overall, breaks = c(-Inf, qs, Inf),
                          labels = c("Critical", "At risk", "Stable",
                                     "Thriving", "Excelling"),
                          include.lowest = TRUE)
  } else {
    dat$risk_band <- cut(dat$idx_overall,
                          breaks = c(-Inf, median(dat$idx_overall, na.rm = TRUE), Inf),
                          labels = c("Below median", "Above median"),
                          include.lowest = TRUE)
  }
  risk_tbl <- dat %>%
    filter(!is.na(risk_band)) %>%
    count(Segment = risk_band) %>%
    mutate(Share = round(100 * n / sum(n), 1))

  low_bands <- c("Critical", "At risk", "Below median")
  pct_at_risk <- round(sum(risk_tbl$Share[risk_tbl$Segment %in% low_bands]), 1)
}

# ---- Sample composition ----------------------------------------------------
comp_rows <- list()
if (sum(!is.na(dat$sector)) >= 20) {
  comp_rows[["Public sector"]]  <- round(100 * mean(dat$sector == "Public", na.rm = TRUE), 1)
  comp_rows[["Private sector"]] <- round(100 * mean(dat$sector == "Private", na.rm = TRUE), 1)
}
if (sum(!is.na(dat$gender)) >= 20) {
  comp_rows[["Female"]] <- round(100 * mean(dat$gender == "Female", na.rm = TRUE), 1)
  comp_rows[["Male"]]   <- round(100 * mean(dat$gender == "Male", na.rm = TRUE), 1)
}
comp_tbl <- NULL
if (length(comp_rows) > 0) {
  comp_tbl <- tibble(Group = names(comp_rows),
                     `Share of sample (%)` = unlist(comp_rows))
}

# ---- Institution league table ----------------------------------------------
uni_tbl <- NULL
if (sum(!is.na(dat$university)) >= 50 && sum(!is.na(dat$idx_overall)) >= 50) {
  uni_tbl <- dat %>%
    filter(!is.na(university), !is.na(idx_overall)) %>%
    group_by(Institution = university) %>%
    summarise(Students = n(), Score = round(mean(idx_overall), 1), .groups = "drop") %>%
    filter(Students >= 5) %>%
    arrange(desc(Score)) %>%
    mutate(Rank = row_number(), Rating = rate_word(Score)) %>%
    select(Rank, Institution, Students, Score, Rating)
}

# ---- Field ranking ----------------------------------------------------------
field_tbl <- NULL
if (sum(!is.na(dat$field)) >= 50 && sum(!is.na(dat$idx_overall)) >= 50) {
  field_tbl <- dat %>%
    filter(!is.na(field), !is.na(idx_overall)) %>%
    group_by(Field = field) %>%
    summarise(Students = n(), Score = round(mean(idx_overall), 1), .groups = "drop") %>%
    filter(Students >= 15) %>%
    arrange(desc(Score)) %>%
    mutate(Rating = rate_word(Score))
}

# ---- Field x Pillar matrix (largest fields, for readability) ----------------
fp_matrix <- NULL
if (!is.null(field_tbl) && nrow(field_tbl) >= 3 && length(PILLARS) >= 2) {
  top_fields <- field_tbl %>% slice_max(Students, n = 8, with_ties = FALSE) %>% pull(Field)
  fp_matrix <- map_dfr(PILLARS, function(v) {
    dat %>%
      filter(field %in% top_fields, !is.na(.data[[v]])) %>%
      group_by(Field = field) %>%
      summarise(Score = round(mean(.data[[v]]), 1), N = n(), .groups = "drop") %>%
      filter(N >= 10) %>%
      mutate(Pillar = LAB[[v]])
  })
  if (nrow(fp_matrix) < 4) fp_matrix <- NULL
}

# ---- Correlation matrix ------------------------------------------------------
CM <- NULL
if (length(PILLARS) >= 2) {
  CM <- cor(dat[, PILLARS], use = "pairwise.complete.obs")
  dimnames(CM) <- list(unname(LAB[PILLARS]), unname(LAB[PILLARS]))
}

# ---- Gender comparison across pillars ----------------------------------------
gender_pillar <- NULL
if (sum(!is.na(dat$gender)) >= 50 && length(PILLARS) >= 2) {
  gender_pillar <- map_dfr(PILLARS, function(v) {
    dat %>%
      filter(!is.na(gender), !is.na(.data[[v]])) %>%
      group_by(Gender = gender) %>%
      summarise(Score = round(mean(.data[[v]]), 1), N = n(), .groups = "drop") %>%
      filter(N >= 25) %>%
      mutate(Pillar = LAB[[v]])
  })
  if (nrow(gender_pillar) < 4) gender_pillar <- NULL
}

# ============================================================================
# ITEM-LEVEL DIAGNOSTICS
# ============================================================================

clean_item_label <- function(col) {
  s <- col
  s <- str_remove(s, "_num$")
  s <- str_remove(s, "^q[0-9]+_[0-9]+_")
  s <- str_remove(s, "^[0-9]+_")
  s <- str_replace_all(s, "_", " ")
  s <- str_squish(s)
  s <- str_to_sentence(s)
  str_trunc(s, 44)
}

item_diag <- NULL
if (nrow(item_map) >= 3) {

  item_diag <- item_map %>%
    mutate(
      Statement = map_chr(var, clean_item_label),
      N         = map_int(var, ~ sum(!is.na(dat[[.x]]))),
      Mean      = map_dbl(var, ~ mean(dat[[.x]], na.rm = TRUE)),
      Pct_agree = map_dbl(var, function(cl) {
        v <- dat[[cl]]; v <- v[!is.na(v)]
        if (!length(v)) NA_real_ else mean(v >= 4)
      }),
      Pct_disagree = map_dbl(var, function(cl) {
        v <- dat[[cl]]; v <- v[!is.na(v)]
        if (!length(v)) NA_real_ else mean(v <= 2)
      }),
      `Net agreement` = round((Pct_agree - Pct_disagree) * 100, 1),
      Performance = round(Mean * 20, 1),
      Importance  = map_dbl(var, function(cl) {
        both <- !is.na(dat[[cl]]) & !is.na(dat$idx_overall)
        if (sum(both) < 20) return(NA_real_)
        tryCatch(cor(dat[[cl]][both], dat$idx_overall[both]),
                 error = function(e) NA_real_)
      })
    ) %>%
    filter(N >= 50, !is.na(`Net agreement`), !is.na(Importance))

  if (nrow(item_diag) < 3) item_diag <- NULL
}

avg_perf <- if (!is.null(item_diag)) mean(item_diag$Performance, na.rm = TRUE) else NA
avg_imp  <- if (!is.null(item_diag)) mean(item_diag$Importance, na.rm = TRUE) else NA

if (!is.null(item_diag)) {
  item_diag <- item_diag %>%
    mutate(Action = case_when(
      Performance <  avg_perf & Importance >= avg_imp ~ "Fix first",
      Performance >= avg_perf & Importance >= avg_imp ~ "Protect strength",
      Performance <  avg_perf & Importance <  avg_imp ~ "Monitor",
      TRUE ~ "Maintain"
    ))
}

top_statement   <- if (!is.null(item_diag)) item_diag$Statement[which.max(item_diag$Performance)] else "N/A"
worst_statement <- if (!is.null(item_diag)) item_diag$Statement[which.min(item_diag$Performance)] else "N/A"
n_fix_first     <- if (!is.null(item_diag)) sum(item_diag$Action == "Fix first") else 0
```

# Executive Summary

## Row {data-height="125"}

### Students Analysed

```{r}
valueBox(comma(total_students), icon = "fa-users", color = INK)
```

### Institutions (5+ responses)

```{r}
valueBox(comma(n_institutions), icon = "fa-university", color = MUTED)
```

### Overall Student Index

```{r}
valueBox(ifelse(is.na(avg_overall), "N/A", paste0(avg_overall, " / 100")),
         icon = "fa-star", color = HIGH)
```

### Survey Statements

```{r}
valueBox(n_items_total, icon = "fa-list-ul", color = MID)
```

## Row {data-height="475"}

### Pillar Scorecard

```{r}
if (!is.null(pillar_tbl)) {
  datatable(pillar_tbl, rownames = FALSE,
            options = list(dom = "t", pageLength = 20, scrollX = TRUE)) %>%
    formatStyle("Index",
                background = styleColorBar(c(0, 100), "#CFE3ED"),
                backgroundSize = "98% 62%",
                backgroundRepeat = "no-repeat",
                backgroundPosition = "left",
                fontWeight = "bold") %>%
    formatStyle("Gap to overall",
                color = styleInterval(0, c(LOW, HIGH)), fontWeight = "bold") %>%
    formatStyle("Rating",
                backgroundColor = styleEqual(
                  c("Poor", "Needs work", "Good", "Very good"),
                  c("#F6D9DA", "#FBEAD2", "#DCEEE7", "#C8E6D9")))
} else {
  datatable(data.frame(Message = "Not enough rating statements to build pillars"),
            rownames = FALSE, options = list(dom = "t"))
}
```

### Student Risk Segmentation

```{r}
if (!is.null(risk_tbl)) {
  seg_colors <- if (all(risk_tbl$Segment %in% names(band_colors_5))) {
    unname(band_colors_5[as.character(risk_tbl$Segment)])
  } else {
    c("#B3202C", "#12715A")
  }
  plot_ly(risk_tbl, labels = ~Segment, values = ~n, type = "pie", hole = 0.55,
          sort = FALSE, direction = "clockwise",
          marker = list(colors = seg_colors, line = list(color = "white", width = 2)),
          textinfo = "label+percent", textposition = "outside",
          hovertemplate = "%{label}<br>%{value} students (%{percent})<extra></extra>") %>%
    layout(showlegend = TRUE,
           legend = list(orientation = "v", x = 1.02, y = 0.5),
           margin = list(t = 30, b = 30, l = 30, r = 30),
           annotations = list(text = paste0("<b>", comma(sum(risk_tbl$n)), "</b><br>students"),
                              showarrow = FALSE, font = list(size = 14)))
} else {
  plotly_empty() %>% layout(title = "Not enough data for risk segmentation")
}
```

## Row {data-height="240"}

### Sample Composition

```{r}
if (!is.null(comp_tbl)) {
  datatable(comp_tbl, rownames = FALSE,
            options = list(dom = "t", pageLength = 10)) %>%
    formatStyle("Share of sample (%)",
                background = styleColorBar(c(0, 100), "#CFE3ED"),
                backgroundSize = "98% 70%",
                backgroundRepeat = "no-repeat",
                backgroundPosition = "left")
} else {
  datatable(data.frame(Message = "Demographic breakdown not available in this dataset"),
            rownames = FALSE, options = list(dom = "t"))
}
```

### Overall Index — How Students Are Spread

```{r}
if (sum(!is.na(dat$idx_overall)) >= 50) {

  hist_vals <- dat$idx_overall[!is.na(dat$idx_overall)]

  plot_ly(x = hist_vals, type = "histogram",
          nbinsx = 28,
          marker = list(color = "#2C5F8A", opacity = 0.85,
                        line = list(color = "white", width = 1)),
          hovertemplate = "Index %{x}<br>%{y} students<extra></extra>") %>%
    layout(
      xaxis = list(title = "Overall student index (0-100)", range = c(0, 100), dtick = 20),
      yaxis = list(title = "Students"),
      margin = list(l = 60, r = 30, t = 20, b = 50),
      shapes = list(
        list(type = "line", x0 = avg_overall, x1 = avg_overall,
             yref = "paper", y0 = 0, y1 = 1,
             line = list(color = LOW, width = 2, dash = "dash"))
      ),
      annotations = list(
        list(x = avg_overall, y = 1.04, xref = "x", yref = "paper",
             text = paste0("<b>Average ", avg_overall, "</b>"),
             showarrow = FALSE, font = list(size = 11, color = LOW))
      ))
} else {
  plotly_empty() %>% layout(title = "Not enough data for this chart")
}
```

# Statement Diagnostics

## Row {data-height="580"}

### Net Agreement by Statement — Ranked Weakest to Strongest

```{r}
if (!is.null(item_diag)) {

  plot_data <- item_diag %>% arrange(`Net agreement`)

  if (nrow(plot_data) > 22) {
    plot_data <- bind_rows(
      plot_data %>% slice_head(n = 11),
      plot_data %>% slice_tail(n = 11)
    ) %>% arrange(`Net agreement`)
  }

  plot_data <- plot_data %>%
    mutate(Statement = factor(Statement, levels = Statement))

  rng <- max(abs(plot_data$`Net agreement`), na.rm = TRUE)
  rng <- ceiling((rng + 14) / 10) * 10

  plot_ly(plot_data,
          x = ~`Net agreement`, y = ~Statement,
          type = "bar", orientation = "h",
          marker = list(color = ~`Net agreement`,
                        colorscale = list(list(0, LOW), list(0.5, "#F2F2F2"), list(1, HIGH)),
                        cmin = -rng, cmax = rng, showscale = FALSE,
                        line = list(color = "white", width = 0.5)),
          text = ~sprintf("%+.0f%%", `Net agreement`),
          textposition = "outside",
          textfont = list(size = 11, color = INK),
          hovertemplate = paste0("<b>%{y}</b><br>Net agreement: %{x:.1f}%",
                                 "<br>Theme: %{customdata}<extra></extra>"),
          customdata = ~Pillar) %>%
    layout(
      xaxis = list(title = "Net agreement  (% who agree minus % who disagree)",
                   range = c(-rng, rng), autorange = FALSE,
                   zeroline = TRUE, zerolinewidth = 2, zerolinecolor = "#333333",
                   ticksuffix = "%"),
      yaxis = list(title = "", automargin = TRUE, tickfont = list(size = 11)),
      margin = list(l = 310, r = 70, t = 20, b = 60),
      bargap = 0.3)
} else {
  plotly_empty() %>% layout(title = "Not enough rating statements for this chart")
}
```

## Row {data-height="470"}

### Priority Matrix — Where to Act First

```{r}
if (!is.null(item_diag)) {

  pal_use <- pillar_palette[names(pillar_palette) %in% unique(item_diag$Pillar)]
  if (length(pal_use) == 0) pal_use <- "Set2"

  plot_ly(item_diag, x = ~Performance, y = ~Importance,
          color = ~Pillar, colors = unname(pal_use),
          type = "scatter", mode = "markers", text = ~Statement,
          marker = list(size = 12, opacity = 0.88,
                        line = list(color = "white", width = 1.2)),
          hovertemplate = paste0("<b>%{text}</b><br>Performance: %{x:.0f}/100",
                                 "<br>Importance: %{y:.2f}<extra></extra>")) %>%
    layout(
      xaxis = list(title = "Performance  (statement score, 0-100)", automargin = TRUE),
      yaxis = list(title = "Importance  (link to overall index)", automargin = TRUE),
      legend = list(orientation = "h", y = -0.25, font = list(size = 10)),
      margin = list(l = 70, r = 30, t = 40, b = 100),
      shapes = list(
        list(type = "line", xref = "x", x0 = avg_perf, x1 = avg_perf,
             yref = "paper", y0 = 0, y1 = 1,
             line = list(dash = "dot", color = MUTED, width = 1.2)),
        list(type = "line", yref = "y", y0 = avg_imp, y1 = avg_imp,
             xref = "paper", x0 = 0, x1 = 1,
             line = list(dash = "dot", color = MUTED, width = 1.2))
      ),
      annotations = list(
        list(x = 0.02, y = 0.98, xref = "paper", yref = "paper",
             text = "<b>FIX FIRST</b>", showarrow = FALSE,
             font = list(size = 10, color = LOW), xanchor = "left"),
        list(x = 0.98, y = 0.98, xref = "paper", yref = "paper",
             text = "<b>PROTECT STRENGTH</b>", showarrow = FALSE,
             font = list(size = 10, color = HIGH), xanchor = "right"),
        list(x = 0.02, y = 0.02, xref = "paper", yref = "paper",
             text = "<b>MONITOR</b>", showarrow = FALSE,
             font = list(size = 10, color = MUTED), xanchor = "left"),
        list(x = 0.98, y = 0.02, xref = "paper", yref = "paper",
             text = "<b>MAINTAIN</b>", showarrow = FALSE,
             font = list(size = 10, color = MUTED), xanchor = "right")
      ))
} else {
  plotly_empty() %>% layout(title = "Not enough rating statements for this chart")
}
```

### Recommended Action per Statement

```{r}
if (!is.null(item_diag)) {
  action_tbl <- item_diag %>%
    select(Statement, Theme = Pillar, Performance, Importance,
           `Net agreement`, Action) %>%
    mutate(Importance = round(Importance, 2)) %>%
    arrange(factor(Action, levels = c("Fix first", "Monitor",
                                       "Protect strength", "Maintain")),
            Performance)

  datatable(action_tbl, rownames = FALSE, filter = "top",
            options = list(pageLength = 8, scrollX = TRUE)) %>%
    formatStyle("Action",
                backgroundColor = styleEqual(
                  c("Fix first", "Protect strength", "Monitor", "Maintain"),
                  c("#F6D9DA", "#C8E6D9", "#FBEAD2", "#DCE7EF")),
                fontWeight = "bold")
} else {
  datatable(data.frame(Message = "Not enough rating statements for this table"),
            rownames = FALSE, options = list(dom = "t"))
}
```

# Institutions

## Row {data-height="560"}

### Highest Ranked Institutions

```{r}
if (!is.null(uni_tbl) && nrow(uni_tbl) > 0) {

  top15 <- uni_tbl %>%
    slice_head(n = 15) %>%
    mutate(Label = str_trunc(Institution, 38)) %>%
    arrange(Score) %>%
    mutate(Label = factor(Label, levels = Label))

  plot_ly(top15,
          x = ~Score, y = ~Label,
          type = "bar", orientation = "h",
          marker = list(color = HIGH, opacity = 0.9,
                        line = list(color = "white", width = 0.6)),
          text = ~sprintf("%.1f  (n=%d)", Score, Students),
          textposition = "outside",
          textfont = list(size = 11, color = INK),
          hovertemplate = paste0("<b>%{y}</b><br>Index: %{x:.1f}/100",
                                 "<br>Respondents: %{customdata}<extra></extra>"),
          customdata = ~Students) %>%
    layout(
      xaxis = list(title = "Overall student index (0-100)",
                   range = c(0, 100), autorange = FALSE,
                   dtick = 20, zeroline = FALSE),
      yaxis = list(title = "", automargin = TRUE, tickfont = list(size = 11)),
      margin = list(l = 280, r = 90, t = 20, b = 60),
      bargap = 0.3)
} else {
  plotly_empty() %>% layout(title = "Institution data not available in this dataset")
}
```

### Lowest Ranked Institutions

```{r}
if (!is.null(uni_tbl) && nrow(uni_tbl) > 0) {

  bot15 <- uni_tbl %>%
    slice_tail(n = 15) %>%
    mutate(Label = str_trunc(Institution, 38)) %>%
    arrange(Score) %>%
    mutate(Label = factor(Label, levels = Label))

  plot_ly(bot15,
          x = ~Score, y = ~Label,
          type = "bar", orientation = "h",
          marker = list(color = LOW, opacity = 0.85,
                        line = list(color = "white", width = 0.6)),
          text = ~sprintf("%.1f  (n=%d)", Score, Students),
          textposition = "outside",
          textfont = list(size = 11, color = INK),
          hovertemplate = paste0("<b>%{y}</b><br>Index: %{x:.1f}/100",
                                 "<br>Respondents: %{customdata}<extra></extra>"),
          customdata = ~Students) %>%
    layout(
      xaxis = list(title = "Overall student index (0-100)",
                   range = c(0, 100), autorange = FALSE,
                   dtick = 20, zeroline = FALSE),
      yaxis = list(title = "", automargin = TRUE, tickfont = list(size = 11)),
      margin = list(l = 280, r = 90, t = 20, b = 60),
      bargap = 0.3)
} else {
  plotly_empty() %>% layout(title = "Institution data not available in this dataset")
}
```

## Row {data-height="440"}

### Complete League Table — Every Institution

```{r}
if (!is.null(uni_tbl) && nrow(uni_tbl) > 0) {
  datatable(uni_tbl, rownames = FALSE, filter = "top",
            extensions = "Buttons",
            options = list(pageLength = 25, scrollX = TRUE, scrollY = "300px",
                           dom = "Bfrtip",
                           buttons = c("copy", "csv", "excel"))) %>%
    formatStyle("Score",
                background = styleColorBar(c(0, 100), "#CFE3ED"),
                backgroundSize = "98% 62%",
                backgroundRepeat = "no-repeat",
                backgroundPosition = "left",
                fontWeight = "bold") %>%
    formatStyle("Rating",
                backgroundColor = styleEqual(
                  c("Poor", "Needs work", "Good", "Very good"),
                  c("#F6D9DA", "#FBEAD2", "#DCEEE7", "#C8E6D9")))
} else {
  datatable(data.frame(Message = "Institution data not available in this dataset"),
            rownames = FALSE, options = list(dom = "t"))
}
```

# Fields of Study

## Row {data-height="480"}

### Overall Index by Field of Study

```{r}
if (!is.null(field_tbl) && nrow(field_tbl) > 0) {

  fplot <- field_tbl %>%
    slice_head(n = 14) %>%
    mutate(Label = str_trunc(Field, 34)) %>%
    arrange(Score) %>%
    mutate(Label = factor(Label, levels = Label))

  plot_ly(fplot,
          x = ~Score, y = ~Label,
          type = "bar", orientation = "h",
          marker = list(color = ~Score,
                        colorscale = list(list(0, LOW), list(0.55, MID), list(1, HIGH)),
                        cmin = 40, cmax = 80, showscale = FALSE,
                        line = list(color = "white", width = 0.6)),
          text = ~sprintf("%.1f  (n=%d)", Score, Students),
          textposition = "outside",
          textfont = list(size = 11, color = INK),
          hovertemplate = paste0("<b>%{y}</b><br>Index: %{x:.1f}/100",
                                 "<br>Students: %{customdata}<extra></extra>"),
          customdata = ~Students) %>%
    layout(
      xaxis = list(title = "Overall student index (0-100)",
                   range = c(0, 100), autorange = FALSE,
                   dtick = 20, zeroline = FALSE),
      yaxis = list(title = "", automargin = TRUE, tickfont = list(size = 11)),
      margin = list(l = 260, r = 90, t = 20, b = 60),
      bargap = 0.3)
} else {
  plotly_empty() %>% layout(title = "Field of study data not available in this dataset")
}
```

## Row {data-height="560"}

### Pillar Strengths and Weaknesses — Eight Largest Fields

```{r}
if (!is.null(fp_matrix)) {

  fp_wide <- fp_matrix %>%
    mutate(Field = str_trunc(Field, 26)) %>%
    select(Field, Pillar, Score) %>%
    pivot_wider(names_from = Pillar, values_from = Score)

  fmat <- as.matrix(fp_wide[, -1, drop = FALSE])
  rownames(fmat) <- fp_wide$Field

  zlo <- floor(min(fmat, na.rm = TRUE)) - 2
  zhi <- ceiling(max(fmat, na.rm = TRUE)) + 2

  plot_ly(x = colnames(fmat), y = rownames(fmat), z = fmat,
          type = "heatmap",
          xgap = 4, ygap = 4,
          colorscale = list(list(0, "#E8A0A6"), list(0.5, "#F4E3B8"), list(1, "#6FAE93")),
          zmin = zlo, zmax = zhi,
          text = round(fmat, 0), texttemplate = "%{text}",
          textfont = list(size = 14, color = INK),
          hovertemplate = "<b>%{y}</b><br>%{x}<br>Index: %{z:.1f}/100<extra></extra>",
          colorbar = list(title = list(text = "Index", side = "right"),
                          len = 0.85, thickness = 14)) %>%
    layout(xaxis = list(title = "", tickangle = -18, automargin = TRUE,
                        tickfont = list(size = 11), side = "bottom"),
           yaxis = list(title = "", automargin = TRUE, tickfont = list(size = 11)),
           margin = list(l = 220, b = 110, t = 20, r = 30))
} else {
  plotly_empty() %>%
    layout(title = "Not enough field-level data to build this comparison")
}
```

# Pillar Analysis

## Row {data-height="560"}

### Pillar Profile

```{r}
if (!is.null(pillar_stats) && nrow(pillar_stats) >= 3) {

  rad <- pillar_stats %>% arrange(Pillar)
  theta_vals <- c(rad$Pillar, rad$Pillar[1])
  r_vals     <- c(rad$Index, rad$Index[1])

  rmin <- max(0, floor((min(rad$Index) - 12) / 10) * 10)

  plot_ly(type = "scatterpolar", mode = "lines+markers",
          r = r_vals, theta = theta_vals,
          fill = "toself",
          fillcolor = "rgba(18,113,90,0.22)",
          line = list(color = HIGH, width = 2.5),
          marker = list(size = 8, color = HIGH),
          hovertemplate = "<b>%{theta}</b><br>Index: %{r:.1f}/100<extra></extra>") %>%
    layout(
      polar = list(
        radialaxis = list(visible = TRUE, range = c(rmin, 100),
                          tickfont = list(size = 10), angle = 90),
        angularaxis = list(tickfont = list(size = 11))
      ),
      showlegend = FALSE,
      margin = list(l = 70, r = 70, t = 50, b = 50))
} else {
  plotly_empty() %>% layout(title = "Need at least three pillars for a profile chart")
}
```

### Typical Score Range per Pillar

```{r}
if (!is.null(pillar_stats)) {

  rngd <- pillar_stats %>%
    arrange(Index) %>%
    mutate(Pillar = factor(Pillar, levels = Pillar))

  pal_r <- pillar_palette[as.character(rngd$Pillar)]
  pal_r[is.na(pal_r)] <- MUTED

  p <- plot_ly()

  for (i in seq_len(nrow(rngd))) {
    p <- p %>%
      add_trace(type = "scatter", mode = "lines",
                x = c(rngd$Q1[i], rngd$Q3[i]),
                y = c(as.character(rngd$Pillar[i]), as.character(rngd$Pillar[i])),
                line = list(color = unname(pal_r[i]), width = 11),
                opacity = 0.42, showlegend = FALSE, hoverinfo = "skip")
  }

  p <- p %>%
    add_trace(type = "scatter", mode = "markers+text",
              x = rngd$Index, y = as.character(rngd$Pillar),
              marker = list(size = 17, color = unname(pal_r),
                            line = list(color = "white", width = 2)),
              text = sprintf("%.0f", rngd$Index),
              textposition = "middle center",
              textfont = list(size = 9, color = "white"),
              showlegend = FALSE,
              hovertemplate = paste0("<b>%{y}</b><br>Average: %{x:.1f}",
                                     "<br>Middle 50%: ", rngd$Q1, " to ", rngd$Q3,
                                     "<extra></extra>")) %>%
    layout(
      xaxis = list(title = "Index score (0-100)", range = c(0, 100), dtick = 20),
      yaxis = list(title = "", automargin = TRUE, tickfont = list(size = 11),
                   categoryorder = "array",
                   categoryarray = as.character(rngd$Pillar)),
      margin = list(l = 210, r = 50, t = 30, b = 60),
      annotations = list(
        list(x = 0.5, y = 1.06, xref = "paper", yref = "paper", showarrow = FALSE,
             text = "Dot = average · Bar = middle 50% of students",
             font = list(size = 11, color = MUTED))
      ))
  p
} else {
  plotly_empty() %>% layout(title = "No pillars available for this chart")
}
```

## Row {data-height="480"}

### How Pillars Move Together

```{r}
if (!is.null(CM)) {

  plot_ly(x = colnames(CM), y = rownames(CM), z = CM,
          type = "heatmap",
          xgap = 4, ygap = 4,
          colorscale = list(list(0, "#F4F7F6"), list(0.5, "#9BC9B8"), list(1, HIGH)),
          zmin = 0, zmax = 1,
          text = round(CM, 2), texttemplate = "%{text}",
          textfont = list(size = 14),
          hovertemplate = "<b>%{y}</b> vs <b>%{x}</b><br>r = %{z:.2f}<extra></extra>",
          colorbar = list(title = list(text = "r", side = "right"),
                          len = 0.85, thickness = 14)) %>%
    layout(xaxis = list(title = "", tickangle = -18, automargin = TRUE,
                        tickfont = list(size = 11)),
           yaxis = list(title = "", automargin = TRUE, tickfont = list(size = 11)),
           margin = list(l = 210, b = 110, t = 20, r = 30))
} else {
  plotly_empty() %>% layout(title = "Need at least two pillars for this chart")
}
```

### Male and Female Students Compared

```{r}
if (!is.null(gender_pillar)) {

  plot_ly(gender_pillar, x = ~Pillar, y = ~Score, color = ~Gender,
          colors = c(Female = "#8C4A83", Male = "#2C5F8A"),
          type = "bar",
          text = ~sprintf("%.0f", Score), textposition = "outside",
          textfont = list(size = 10, color = INK),
          hovertemplate = "<b>%{x}</b><br>%{data.name}: %{y:.1f}/100<extra></extra>") %>%
    layout(barmode = "group",
           yaxis = list(title = "Index score (0-100)", range = c(0, 100), dtick = 20),
           xaxis = list(title = "", tickangle = -18, automargin = TRUE,
                        tickfont = list(size = 10)),
           legend = list(orientation = "h", y = 1.12),
           margin = list(l = 60, r = 30, t = 40, b = 110))
} else {
  plotly_empty() %>% layout(title = "Gender comparison not available in this dataset")
}
```

# Key Insights

## Row {data-height="125"}

### Strongest Pillar

```{r}
valueBox(ifelse(is.na(best_score), "N/A", paste0(best_pillar, " — ", best_score)),
         icon = "fa-arrow-up", color = HIGH)
```

### Weakest Pillar

```{r}
valueBox(ifelse(is.na(worst_score), "N/A", paste0(worst_pillar, " — ", worst_score)),
         icon = "fa-arrow-down", color = LOW)
```

### Students Below Average

```{r}
valueBox(ifelse(is.na(pct_at_risk), "N/A", paste0(pct_at_risk, "%")),
         icon = "fa-exclamation-triangle", color = MID)
```

### Statements to Fix First

```{r}
valueBox(n_fix_first, icon = "fa-wrench", color = INK)
```

## Row {data-height="520"}

### Ten Statements With the Greatest Influence on Overall Experience

```{r}
if (!is.null(item_diag)) {

  imp_top <- item_diag %>%
    arrange(desc(Importance)) %>%
    slice_head(n = 10) %>%
    arrange(Importance) %>%
    mutate(Statement = factor(Statement, levels = Statement))

  plot_ly(imp_top,
          x = ~Importance, y = ~Statement,
          type = "bar", orientation = "h",
          marker = list(color = ~Performance,
                        colorscale = list(list(0, LOW), list(0.5, MID), list(1, HIGH)),
                        cmin = 40, cmax = 80,
                        colorbar = list(title = list(text = "Score", side = "right"),
                                        len = 0.85, thickness = 13),
                        line = list(color = "white", width = 0.6)),
          text = ~sprintf("%.2f", Importance),
          textposition = "outside",
          textfont = list(size = 11, color = INK),
          hovertemplate = paste0("<b>%{y}</b><br>Influence: %{x:.2f}",
                                 "<br>Its own score: %{marker.color:.0f}/100<extra></extra>")) %>%
    layout(
      xaxis = list(title = "Influence on the overall index (correlation)",
                   range = c(0, max(imp_top$Importance) * 1.25), zeroline = FALSE),
      yaxis = list(title = "", automargin = TRUE, tickfont = list(size = 11)),
      margin = list(l = 310, r = 60, t = 20, b = 60),
      bargap = 0.3)
} else {
  plotly_empty() %>% layout(title = "Not enough rating statements for this chart")
}
```

## Row {data-height="475"}

### Where Each Pillar Sits Against the Overall Average

```{r}
if (!is.null(pillar_stats)) {

  gap_d <- pillar_stats %>%
    arrange(`Gap to overall`) %>%
    mutate(Pillar = factor(Pillar, levels = Pillar),
           Direction = ifelse(`Gap to overall` >= 0, "Above average", "Below average"))

  lim <- max(abs(gap_d$`Gap to overall`), na.rm = TRUE)
  lim <- ceiling((lim + 2) / 2) * 2

  plot_ly(gap_d, x = ~`Gap to overall`, y = ~Pillar,
          type = "bar", orientation = "h",
          color = ~Direction,
          colors = c(`Above average` = HIGH, `Below average` = LOW),
          text = ~sprintf("%+.1f", `Gap to overall`),
          textposition = "outside",
          textfont = list(size = 11, color = INK),
          hovertemplate = "<b>%{y}</b><br>Gap: %{x:+.1f} points<extra></extra>") %>%
    layout(
      xaxis = list(title = "Points above or below the overall index",
                   range = c(-lim, lim), zeroline = TRUE,
                   zerolinewidth = 2, zerolinecolor = "#333333"),
      yaxis = list(title = "", automargin = TRUE, tickfont = list(size = 11)),
      legend = list(orientation = "h", y = 1.15),
      margin = list(l = 210, r = 60, t = 40, b = 60),
      bargap = 0.35)
} else {
  plotly_empty() %>% layout(title = "No pillars available for this chart")
}
```

### Sample Size Behind Every Field

```{r}
if (!is.null(field_tbl) && nrow(field_tbl) > 0) {

  fn <- field_tbl %>%
    arrange(desc(Students)) %>%
    slice_head(n = 12) %>%
    arrange(Students) %>%
    mutate(Label = str_trunc(Field, 30),
           Label = factor(Label, levels = Label))

  plot_ly(fn, x = ~Students, y = ~Label,
          type = "bar", orientation = "h",
          marker = list(color = "#2C5F8A", opacity = 0.88,
                        line = list(color = "white", width = 0.6)),
          text = ~comma(Students), textposition = "outside",
          textfont = list(size = 11, color = INK),
          hovertemplate = "<b>%{y}</b><br>%{x} students<extra></extra>") %>%
    layout(
      xaxis = list(title = "Number of students", zeroline = FALSE,
                   range = c(0, max(fn$Students) * 1.22)),
      yaxis = list(title = "", automargin = TRUE, tickfont = list(size = 11)),
      margin = list(l = 240, r = 60, t = 20, b = 60),
      bargap = 0.3)
} else {
  plotly_empty() %>% layout(title = "Field of study data not available in this dataset")
}
```

# Data Explorer

## Row

### Browse, Sort, Filter and Export the Full Dataset

```{r}
display_cols <- c("sector", "gender", "age", "university", "field",
                   PILLARS, "idx_overall")
display_cols <- display_cols[display_cols %in% names(dat)]

explorer_data <- dat %>% select(all_of(display_cols))

rename_vec <- c(sector = "Sector", gender = "Gender", age = "Age",
                university = "Institution", field = "Field of study",
                idx_overall = "Overall index")
for (v in PILLARS) rename_vec[v] <- LAB[[v]]

names(explorer_data) <- unname(ifelse(names(explorer_data) %in% names(rename_vec),
                                       rename_vec[names(explorer_data)],
                                       names(explorer_data)))

num_display <- names(explorer_data)[sapply(explorer_data, is.numeric)]

dt <- datatable(explorer_data, rownames = FALSE, filter = "top",
                extensions = "Buttons",
                options = list(pageLength = 15, scrollX = TRUE,
                               dom = "Bfrtip",
                               buttons = c("copy", "csv", "excel")))

if (length(num_display) > 0) {
  dt <- dt %>% formatRound(columns = num_display, digits = 1)
}

dt
```
