---
title: "University Student Survey — Executive Dashboard"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    theme: flatly
---

<style>
body {
  font-family: -apple-system, "Segoe UI", Helvetica, Arial, sans-serif;
}
.navbar {
  font-weight: 600;
  letter-spacing: 0.02em;
}
.chart-title {
  font-weight: 700 !important;
  font-size: 15px !important;
  letter-spacing: 0.01em;
  color: #14202E !important;
  border-bottom: 2px solid #E4E8EC;
}
.value-box .value {
  font-weight: 700 !important;
}
.value-box .caption {
  font-weight: 500 !important;
  opacity: 0.92;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# ---- Auto-install any missing packages -----------------------------------
required_pkgs <- c("flexdashboard", "tidyverse", "readxl", "janitor",
                    "plotly", "DT", "scales")
for (pkg in required_pkgs) {
  if (!requireNamespace(pkg, quietly = TRUE)) install.packages(pkg)
}

library(flexdashboard)
library(tidyverse)
library(readxl)
library(janitor)
library(plotly)
library(DT)
library(scales)

set.seed(2026)

# ================================================================ PALETTE ===
INK   <- "#14202E"
MUTED <- "#7A8794"
LOW   <- "#B3202C"
MID   <- "#E8B33C"
HIGH  <- "#12715A"
sect_colors <- c(Public = "#2C5F8A", Private = "#D9772E")
band_colors_5 <- c("Critical" = "#8C1720", "At risk" = "#C75B3F",
                     "Stable" = "#D9B23C", "Thriving" = "#5FA88E",
                     "Excelling" = "#12715A")

# ============================================================================
# LOAD & CLEAN DATA
# (Book1.xlsx must sit in the same folder as this .Rmd file when you knit it)
# ============================================================================

dat <- read_excel("Book1.xlsx") %>% clean_names()

dat <- dat %>%
  select(-matches("name|respondent_id|timestamp|email|phone|enrollment|roll_no|student_id",
                  ignore.case = TRUE))

dat <- dat[rowMeans(is.na(dat)) <= 0.6, ]
dat <- dat[!duplicated(dat), ]

# ---- Detect key demographic columns ---------------------------------------
sector_col <- grep("sector|institution_type", names(dat), value = TRUE)[1]
gender_col <- grep("gender|sex", names(dat), value = TRUE)[1]
univ_col   <- grep("university_reported|university|institution", names(dat), value = TRUE)[1]
field_col  <- grep("field_group|discipline|faculty|program|field", names(dat), value = TRUE)[1]
age_col    <- grep("^q.*age|^age", names(dat), value = TRUE)[1]

if (!is.na(sector_col)) {
  s <- tolower(trimws(as.character(dat[[sector_col]])))
  s[grepl("govt|government|public|federal", s)] <- "Public"
  s[grepl("pvt|private", s)] <- "Private"
  s[!s %in% c("Public", "Private")] <- NA
  dat$sector <- s
} else {
  dat$sector <- NA
}

if (!is.na(gender_col)) {
  g <- tolower(trimws(as.character(dat[[gender_col]])))
  g[g %in% c("m", "male")] <- "Male"
  g[g %in% c("f", "female")] <- "Female"
  g[!g %in% c("Male", "Female")] <- NA
  dat$gender <- g
} else {
  dat$gender <- NA
}

if (!is.na(univ_col)) {
  dat$university <- str_to_title(str_squish(as.character(dat[[univ_col]])))
  dat$university[dat$university %in% c("", "Na", "N/A")] <- NA
} else {
  dat$university <- NA
}

if (!is.na(field_col)) {
  dat$field <- str_to_title(str_squish(as.character(dat[[field_col]])))
  dat$field[dat$field %in% c("", "Na", "N/A")] <- NA
} else {
  dat$field <- NA
}

if (!is.na(age_col)) {
  dat$age <- suppressWarnings(as.numeric(dat[[age_col]]))
  dat$age[dat$age < 15 | dat$age > 60] <- NA
} else {
  dat$age <- NA
}

# ---- Likert conversion -----------------------------------------------------
lk <- c("strongly disagree" = 1, "disagree" = 2, "neutral" = 3,
        "neither agree nor disagree" = 3, "agree" = 4, "strongly agree" = 5,
        "very dissatisfied" = 1, "dissatisfied" = 2, "satisfied" = 4,
        "very satisfied" = 5, "very poor" = 1, "poor" = 2, "average" = 3,
        "good" = 4, "very good" = 5, "excellent" = 5, "fair" = 3,
        "never" = 1, "rarely" = 2, "sometimes" = 3, "often" = 4, "always" = 5,
        "not at all" = 1, "slightly" = 2, "moderately" = 3, "very" = 4,
        "extremely" = 5, "1" = 1, "2" = 2, "3" = 3, "4" = 4, "5" = 5)

to_num <- function(x) {
  v <- if (is.numeric(x)) x else unname(lk[tolower(trimws(as.character(x)))])
  v <- as.numeric(v)
  v[!is.na(v) & (v < 1 | v > 5)] <- NA
  v
}

raw_likert <- grep("^q[0-9]+_[0-9]+_", names(dat), value = TRUE)
raw_likert <- raw_likert[!grepl("_num$", raw_likert)]
for (cl in raw_likert) {
  cv <- to_num(dat[[cl]])
  if (sum(!is.na(cv)) >= 30) dat[[paste0(cl, "_num")]] <- cv
}

num_cols <- grep("_num$", names(dat), value = TRUE)

# ---- Build pillar indices ---------------------------------------------------
blocks <- split(num_cols, str_extract(num_cols, "^q[0-9]+_[0-9]+"))
blocks <- blocks[sapply(blocks, length) >= 2]

label_for <- function(cols, pref) {
  stem <- cols %>% str_remove("_num$") %>% str_remove(paste0("^", pref, "_[0-9]+_"))
  w <- unlist(str_split(stem, "_"))
  w <- w[nchar(w) > 3]
  if (!length(w)) return(toupper(pref))
  str_to_title(names(sort(table(w), decreasing = TRUE))[1])
}

row_index <- function(d, cols) {
  m <- suppressWarnings(apply(as.matrix(d[, cols, drop = FALSE]), 2, as.numeric))
  if (is.null(dim(m))) m <- matrix(m, ncol = 1)
  o <- rowMeans(m, na.rm = TRUE) * 20
  o[rowSums(!is.na(m)) == 0] <- NA_real_
  pmin(100, pmax(0, o))
}

idx_tbl <- tibble(prefix = names(blocks)) %>%
  mutate(label = map2_chr(blocks, prefix, label_for), var = paste0("idx_", prefix))

for (i in seq_len(nrow(idx_tbl))) {
  dat[[idx_tbl$var[i]]] <- row_index(dat, blocks[[idx_tbl$prefix[i]]])
}

dat$idx_overall <- row_index(dat, num_cols)
idx_tbl <- bind_rows(idx_tbl, tibble(prefix = "all", label = "Overall", var = "idx_overall"))
idx_tbl <- idx_tbl[sapply(idx_tbl$var, function(v) sum(!is.na(dat[[v]]))) >= 50, ]
idx_tbl <- idx_tbl[order(-sapply(idx_tbl$var, function(v) mean(dat[[v]], na.rm = TRUE))), ]

IDX <- idx_tbl$var
LAB <- setNames(idx_tbl$label, idx_tbl$var)
PILLARS <- setdiff(IDX, "idx_overall")

rate_word <- function(m) {
  case_when(m >= 80 ~ "Very good", m >= 60 ~ "Good",
            m >= 40 ~ "Needs work", TRUE ~ "Poor")
}

# ---- Summary numbers used in value boxes ------------------------------------
total_students <- nrow(dat)
avg_overall     <- if (sum(!is.na(dat$idx_overall)) > 0) round(mean(dat$idx_overall, na.rm = TRUE), 1) else NA
n_indices       <- length(PILLARS)
n_institutions  <- if (sum(!is.na(dat$university)) > 0) n_distinct(dat$university, na.rm = TRUE) else NA

# ---- Pillar summary table (Index / Gap to overall / Rating) ----------------
pillar_tbl <- NULL
if (length(PILLARS) > 0) {
  pillar_tbl <- map_dfr(PILLARS, function(v) {
    x <- dat[[v]][!is.na(dat[[v]])]
    tibble(Pillar = LAB[[v]], N = length(x), Index = round(mean(x), 1))
  }) %>%
    mutate(`Gap to overall` = round(Index - avg_overall, 1),
           Rating = rate_word(Index)) %>%
    arrange(desc(Index))
}

# ---- Risk segmentation (adaptive 5-band, falls back safely on tied data) ---
risk_tbl <- NULL
if (sum(!is.na(dat$idx_overall)) >= 50) {
  qs <- quantile(dat$idx_overall, probs = c(.2, .4, .6, .8), na.rm = TRUE, type = 7)
  qs <- unique(qs)
  if (length(qs) == 4) {
    breaks_use <- c(-Inf, qs, Inf)
    labels_use <- c("Critical", "At risk", "Stable", "Thriving", "Excelling")
  } else {
    breaks_use <- c(-Inf, median(dat$idx_overall, na.rm = TRUE), Inf)
    labels_use <- c("Below median", "Above median")
  }
  dat$risk_band <- cut(dat$idx_overall, breaks = breaks_use, labels = labels_use,
                        include.lowest = TRUE)
  risk_tbl <- dat %>%
    filter(!is.na(risk_band)) %>%
    count(Segment = risk_band) %>%
    mutate(Share = round(100 * n / sum(n), 1))
}

# ---- Sample composition ------------------------------------------------------
comp_rows <- list()
if (sum(!is.na(dat$sector)) >= 20) {
  comp_rows[["Public sector"]] <- round(100 * mean(dat$sector == "Public", na.rm = TRUE), 1)
  comp_rows[["Private sector"]] <- round(100 * mean(dat$sector == "Private", na.rm = TRUE), 1)
}
if (sum(!is.na(dat$gender)) >= 20) {
  comp_rows[["Female"]] <- round(100 * mean(dat$gender == "Female", na.rm = TRUE), 1)
  comp_rows[["Male"]] <- round(100 * mean(dat$gender == "Male", na.rm = TRUE), 1)
}
comp_tbl <- NULL
if (length(comp_rows) > 0) {
  comp_tbl <- tibble(Group = names(comp_rows), `Share of sample (%)` = unlist(comp_rows))
}

# ---- Institution ranking table ------------------------------------------------
uni_tbl <- NULL
if (sum(!is.na(dat$university)) >= 50 && sum(!is.na(dat$idx_overall)) >= 50) {
  uni_tbl <- dat %>%
    filter(!is.na(university), !is.na(idx_overall)) %>%
    group_by(University = university) %>%
    summarise(Students = n(), `Average score` = round(mean(idx_overall), 1),
              .groups = "drop") %>%
    filter(Students >= 5) %>%
    arrange(desc(`Average score`)) %>%
    mutate(Rating = rate_word(`Average score`))
}

# ---- Field ranking table -------------------------------------------------------
field_tbl <- NULL
if (sum(!is.na(dat$field)) >= 50 && sum(!is.na(dat$idx_overall)) >= 50) {
  field_tbl <- dat %>%
    filter(!is.na(field), !is.na(idx_overall)) %>%
    group_by(Field = field) %>%
    summarise(Students = n(), `Average score` = round(mean(idx_overall), 1),
              .groups = "drop") %>%
    filter(Students >= 10) %>%
    arrange(desc(`Average score`)) %>%
    mutate(Rating = rate_word(`Average score`))
}

# ---- Sector x pillar table -------------------------------------------------------
sector_idx_tbl <- NULL
if (sum(!is.na(dat$sector)) >= 50 && length(PILLARS) >= 1) {
  sector_idx_tbl <- map_dfr(PILLARS, function(v) {
    dat %>%
      filter(!is.na(sector), !is.na(.data[[v]])) %>%
      group_by(Sector = sector) %>%
      summarise(Mean = round(mean(.data[[v]]), 1), .groups = "drop") %>%
      mutate(Pillar = LAB[[v]])
  })
}

# ---- Correlation matrix (vectorised) -----------------------------------------
CM <- NULL
if (length(PILLARS) >= 2) {
  CM <- cor(dat[, PILLARS], use = "pairwise.complete.obs")
  dimnames(CM) <- list(unname(LAB[PILLARS]), unname(LAB[PILLARS]))
}

# ============================================================================
# ITEM-LEVEL DIAGNOSTICS (every individual statement, not just the pillar)
# Mirrors a "diverging agreement profile" + "importance vs performance" view.
# Fully derived from num_cols — no manual pairwise loops, so nothing here can
# reproduce the earlier indexing bug: each item is scored independently
# against the fixed idx_overall vector, never against another item.
# ============================================================================

item_diag <- NULL
if (length(num_cols) >= 5) {

  item_label_of <- function(col) {
    stem <- col %>% str_remove("_num$") %>% str_remove("^q[0-9]+_[0-9]+_")
    stem <- str_replace_all(stem, "_", " ")
    lbl <- str_to_sentence(stem)
    if (nchar(lbl) < 2) return(col)
    lbl
  }

  item_diag <- tibble(var = num_cols) %>%
    mutate(
      prefix  = str_extract(var, "^q[0-9]+_[0-9]+"),
      idx_var = paste0("idx_", prefix),
      Section = unname(LAB[idx_var]),
      Section = ifelse(is.na(Section), "Other", Section),
      Label   = map_chr(var, item_label_of),
      N       = map_int(var, ~ sum(!is.na(dat[[.x]]))),
      Mean    = map_dbl(var, ~ mean(dat[[.x]], na.rm = TRUE)),
      Pct_agree = map_dbl(var, function(cl) {
        v <- dat[[cl]]; v <- v[!is.na(v)]
        if (length(v) == 0) NA_real_ else mean(v >= 4)
      }),
      Pct_disagree = map_dbl(var, function(cl) {
        v <- dat[[cl]]; v <- v[!is.na(v)]
        if (length(v) == 0) NA_real_ else mean(v <= 2)
      }),
      Net_agreement = round((Pct_agree - Pct_disagree) * 100, 1),
      Performance   = round(Mean * 20, 1),
      Importance    = map_dbl(var, function(cl) {
        both <- !is.na(dat[[cl]]) & !is.na(dat$idx_overall)
        if (sum(both) < 20) return(NA_real_)
        tryCatch(cor(dat[[cl]][both], dat$idx_overall[both]),
                 error = function(e) NA_real_)
      })
    ) %>%
    filter(N >= 30, !is.na(Net_agreement))

  if (nrow(item_diag) == 0) item_diag <- NULL
}

avg_perf <- if (!is.null(item_diag)) mean(item_diag$Performance, na.rm = TRUE) else NA
avg_imp  <- if (!is.null(item_diag)) mean(item_diag$Importance, na.rm = TRUE) else NA

if (!is.null(item_diag)) {
  item_diag <- item_diag %>%
    mutate(Quadrant = case_when(
      Performance <  avg_perf & Importance >= avg_imp ~ "Fix first",
      Performance >= avg_perf & Importance >= avg_imp ~ "Protect strength",
      Performance <  avg_perf & Importance <  avg_imp ~ "Monitor",
      TRUE ~ "Maintain"
    ))
}
```

Executive Summary
=======================================================================

Row {data-height=130}
-----------------------------------------------------------------------

### Total Students

```{r}
valueBox(comma(total_students), icon = "fa-users", color = INK)
```

### Institutions Represented

```{r}
valueBox(ifelse(is.na(n_institutions), "N/A", comma(n_institutions)),
         icon = "fa-university", color = MUTED)
```

### Overall Student Index

```{r}
valueBox(ifelse(is.na(avg_overall), "N/A", paste0(avg_overall, " / 100")),
         icon = "fa-star", color = HIGH)
```

### Pillars Measured

```{r}
valueBox(n_indices, icon = "fa-list", color = MID)
```

Row {data-height=430}
-----------------------------------------------------------------------

### Pillar Scorecard — Index, Gap to Overall, Rating

```{r}
if (!is.null(pillar_tbl)) {
  datatable(pillar_tbl, rownames = FALSE,
            options = list(dom = "t", pageLength = 20, scrollX = TRUE)) %>%
    formatStyle("Gap to overall",
                color = styleInterval(0, c(LOW, HIGH)),
                fontWeight = "bold") %>%
    formatStyle("Rating",
                backgroundColor = styleEqual(
                  c("Poor", "Needs work", "Good", "Very good"),
                  c("#F6D9DA", "#FBEAD2", "#DCEEE7", "#C8E6D9")))
} else {
  datatable(data.frame(Message = "Not enough item-level data to build pillars"),
            rownames = FALSE, options = list(dom = "t"))
}
```

### Student Risk Segmentation

```{r}
if (!is.null(risk_tbl)) {
  seg_colors <- if (all(risk_tbl$Segment %in% names(band_colors_5))) {
    unname(band_colors_5[as.character(risk_tbl$Segment)])
  } else {
    c("#B3202C", "#12715A")
  }
  plot_ly(risk_tbl, labels = ~Segment, values = ~n, type = "pie", hole = 0.55,
          marker = list(colors = seg_colors, line = list(color = "white", width = 2)),
          textinfo = "label+percent",
          hovertemplate = "%{label}<br>%{value} students (%{percent})<extra></extra>") %>%
    layout(showlegend = TRUE,
           annotations = list(text = paste0(sum(risk_tbl$n), "\nstudents"),
                              showarrow = FALSE, font = list(size = 15)))
} else {
  plotly_empty() %>% layout(title = "Not enough data for risk segmentation")
}
```

Row {data-height=250}
-----------------------------------------------------------------------

### Sample Composition

```{r}
if (!is.null(comp_tbl)) {
  datatable(comp_tbl, rownames = FALSE,
            options = list(dom = "t", pageLength = 10)) %>%
    formatStyle("Share of sample (%)",
                background = styleColorBar(c(0, 100), "#CFE3ED"),
                backgroundSize = "98% 70%",
                backgroundRepeat = "no-repeat",
                backgroundPosition = "left")
} else {
  datatable(data.frame(Message = "Demographic breakdown not available in this dataset"),
            rownames = FALSE, options = list(dom = "t"))
}
```

### How to Read This Page

Every pillar is the mean of its member survey items, rescaled to run from
20 to 100 so pillars are directly comparable — higher is always better.
**Gap to overall** shows how far a pillar sits from the overall student
index; a negative gap marks a pillar dragging the average down.

Students are split into five **risk segments** using the quantiles of
their own overall index, so segment sizes stay meaningful whatever the
underlying distribution looks like. **Critical** and **At risk** together
are the population worth reaching first with any intervention.

Statement Diagnostics
=======================================================================

Row
-----------------------------------------------------------------------

### Diverging Agreement Profile — Ranked Worst to Best

```{r, fig.height=8}
if (!is.null(item_diag)) {

  plot_data <- item_diag %>% arrange(Net_agreement)
  if (nrow(plot_data) > 30) {
    plot_data <- bind_rows(
      plot_data %>% slice_head(n = 15),
      plot_data %>% slice_tail(n = 15)
    ) %>% arrange(Net_agreement)
  }

  plot_ly(plot_data, x = ~Net_agreement, y = ~reorder(Label, Net_agreement),
          type = "bar", orientation = "h",
          marker = list(color = ~Net_agreement,
                        colorscale = list(c(0, LOW), c(0.5, "#F2F2F2"), c(1, HIGH)),
                        cmin = -60, cmax = 60, showscale = FALSE),
          text = ~sprintf("%+.1f", Net_agreement), textposition = "outside",
          hovertemplate = paste0("%{y}<br>Net agreement: %{x:.1f}",
                                 "<br>Section: %{customdata}<extra></extra>"),
          customdata = ~Section) %>%
    layout(xaxis = list(title = "Net agreement (% agree − % disagree)", zeroline = TRUE,
                        zerolinewidth = 2, zerolinecolor = MUTED),
           yaxis = list(title = ""),
           margin = list(l = 260))
} else {
  plotly_empty() %>% layout(title = "Not enough item-level data for this chart")
}
```

Row
-----------------------------------------------------------------------

### Priority Matrix — Performance vs Importance

```{r}
if (!is.null(item_diag)) {

  plot_ly(item_diag, x = ~Performance, y = ~Importance, color = ~Section,
          colors = "Set2", type = "scatter", mode = "markers", text = ~Label,
          marker = list(size = 10, opacity = 0.8, line = list(color = "white", width = 0.6)),
          hovertemplate = "%{text}<br>Performance: %{x:.0f}<br>Importance: %{y:.2f}<extra></extra>") %>%
    layout(
      xaxis = list(title = "Performance (0-100 index score)"),
      yaxis = list(title = "Importance (correlation with overall index)"),
      shapes = list(
        list(type = "line", xref = "x", x0 = avg_perf, x1 = avg_perf,
             yref = "paper", y0 = 0, y1 = 1,
             line = list(dash = "dot", color = MUTED, width = 1)),
        list(type = "line", yref = "y", y0 = avg_imp, y1 = avg_imp,
             xref = "paper", x0 = 0, x1 = 1,
             line = list(dash = "dot", color = MUTED, width = 1))
      ))
} else {
  plotly_empty() %>% layout(title = "Not enough item-level data for this chart")
}
```

### Recommended Action per Statement

```{r}
if (!is.null(item_diag)) {
  action_tbl <- item_diag %>%
    select(Statement = Label, Section, Performance, Importance, Quadrant) %>%
    mutate(Importance = round(Importance, 2)) %>%
    arrange(Quadrant, Performance)

  datatable(action_tbl, rownames = FALSE, filter = "top",
            options = list(pageLength = 10, scrollX = TRUE)) %>%
    formatStyle("Quadrant",
                backgroundColor = styleEqual(
                  c("Fix first", "Protect strength", "Monitor", "Maintain"),
                  c("#F6D9DA", "#C8E6D9", "#FBEAD2", "#DCE7EF")))
} else {
  datatable(data.frame(Message = "Not enough item-level data for this table"),
            rownames = FALSE, options = list(dom = "t"))
}
```

Institutions
=======================================================================

Row
-----------------------------------------------------------------------

### Top 15 Institutions by Overall Index

```{r}
if (!is.null(uni_tbl) && nrow(uni_tbl) > 0) {
  top15 <- uni_tbl %>% slice_head(n = 15) %>% arrange(`Average score`)

  plot_ly(top15, x = ~`Average score`, y = ~reorder(University, `Average score`),
          type = "bar", orientation = "h",
          marker = list(color = ~`Average score`,
                        colorscale = list(c(0, LOW), c(0.5, MID), c(1, HIGH)),
                        cmin = 0, cmax = 100, showscale = FALSE),
          text = ~sprintf("%.0f (n=%d)", `Average score`, Students),
          textposition = "outside",
          hovertemplate = "%{y}<br>Score: %{x:.1f}<extra></extra>") %>%
    layout(xaxis = list(title = "Average score (0-100)", range = c(0, 115)),
           yaxis = list(title = ""),
           margin = list(l = 220))
} else {
  plotly_empty() %>% layout(title = "Institution data not available in this dataset")
}
```

Row
-----------------------------------------------------------------------

### Full Institution League Table

```{r}
if (!is.null(uni_tbl) && nrow(uni_tbl) > 0) {
  datatable(uni_tbl, rownames = FALSE, filter = "top",
            options = list(pageLength = 12, scrollX = TRUE)) %>%
    formatStyle("Rating",
                backgroundColor = styleEqual(
                  c("Poor", "Needs work", "Good", "Very good"),
                  c("#F6D9DA", "#FBEAD2", "#DCEEE7", "#C8E6D9")))
} else {
  datatable(data.frame(Message = "Institution data not available in this dataset"),
            rownames = FALSE, options = list(dom = "t"))
}
```

Fields & Sectors
=======================================================================

Row
-----------------------------------------------------------------------

### Satisfaction by Field of Study

```{r}
if (!is.null(field_tbl) && nrow(field_tbl) > 0) {
  fplot <- field_tbl %>% arrange(`Average score`)

  plot_ly(fplot, x = ~`Average score`, y = ~reorder(Field, `Average score`),
          type = "bar", orientation = "h",
          marker = list(color = ~`Average score`,
                        colorscale = list(c(0, LOW), c(0.5, MID), c(1, HIGH)),
                        cmin = 0, cmax = 100, showscale = FALSE),
          text = ~sprintf("%.0f (n=%d)", `Average score`, Students),
          textposition = "outside",
          hovertemplate = "%{y}<br>Score: %{x:.1f}<extra></extra>") %>%
    layout(xaxis = list(title = "Average score (0-100)", range = c(0, 115)),
           yaxis = list(title = ""),
           margin = list(l = 200))
} else {
  plotly_empty() %>% layout(title = "Field of study data not available in this dataset")
}
```

### Public vs Private — Every Pillar Compared

```{r}
if (!is.null(sector_idx_tbl) && nrow(sector_idx_tbl) > 0) {
  plot_ly(sector_idx_tbl, x = ~Pillar, y = ~Mean, color = ~Sector, type = "bar",
          colors = sect_colors,
          text = ~sprintf("%.0f", Mean), textposition = "outside",
          hovertemplate = "%{x}<br>%{data.name}: %{y:.1f}<extra></extra>") %>%
    layout(barmode = "group",
           yaxis = list(title = "Average score (0-100)", range = c(0, 115)),
           xaxis = list(title = ""),
           legend = list(orientation = "h", y = 1.1))
} else {
  plotly_empty() %>% layout(title = "Sector data not available in this dataset")
}
```

Relationships
=======================================================================

Row
-----------------------------------------------------------------------

### Scatter: Two Strongest Pillars

```{r}
if (length(PILLARS) >= 2) {

  sc_data <- dat %>%
    select(x = all_of(PILLARS[1]), y = all_of(PILLARS[2]), sector) %>%
    mutate(x = suppressWarnings(as.numeric(x)),
           y = suppressWarnings(as.numeric(y))) %>%
    filter(!is.na(x), !is.na(y))

  if (nrow(sc_data) >= 20) {

    sc_data$sector_display <- ifelse(is.na(sc_data$sector), "Unknown", sc_data$sector)
    point_colors <- c(sect_colors, Unknown = "#AAAAAA")

    plot_ly(sc_data, x = ~x, y = ~y, color = ~sector_display,
            colors = point_colors, type = "scatter", mode = "markers",
            marker = list(size = 8, opacity = 0.65,
                          line = list(color = "white", width = 0.5)),
            hovertemplate = paste0(LAB[[PILLARS[1]]], ": %{x:.0f}<br>",
                                   LAB[[PILLARS[2]]], ": %{y:.0f}<extra></extra>")) %>%
      layout(xaxis = list(title = LAB[[PILLARS[1]]], range = c(0, 100)),
             yaxis = list(title = LAB[[PILLARS[2]]], range = c(0, 100)),
             legend = list(title = list(text = "Sector")))
  } else {
    plotly_empty() %>% layout(title = "Not enough overlapping data for this scatter plot")
  }
} else {
  plotly_empty() %>% layout(title = "Need at least two pillars for a scatter plot")
}
```

### Correlation Heatmap

```{r}
if (!is.null(CM)) {

  plot_ly(x = colnames(CM), y = rownames(CM), z = CM, type = "heatmap",
          colorscale = list(c(0, "#2166AC"), c(0.5, "#F7F8FA"), c(1, "#B2182B")),
          zmin = -1, zmax = 1,
          text = round(CM, 2), texttemplate = "%{text}",
          hovertemplate = "%{y} vs %{x}<br>r = %{z:.2f}<extra></extra>") %>%
    layout(xaxis = list(title = "", tickangle = -35),
           yaxis = list(title = ""),
           margin = list(l = 160, b = 120))
} else {
  plotly_empty() %>% layout(title = "Need at least two pillars for a correlation heatmap")
}
```

Data Explorer
=======================================================================

Row
-----------------------------------------------------------------------

### Browse, Sort, Filter, and Search the Cleaned Data

```{r}
display_cols <- c("sector", "gender", "age", "university", "field", PILLARS, "idx_overall")
display_cols <- display_cols[display_cols %in% names(dat)]

explorer_data <- dat %>%
  select(all_of(display_cols)) %>%
  rename_with(~ LAB[.x], .cols = any_of(c(PILLARS, "idx_overall")))

num_display <- names(explorer_data)[sapply(explorer_data, is.numeric)]

dt <- datatable(explorer_data, rownames = FALSE, filter = "top",
                options = list(pageLength = 15, scrollX = TRUE,
                               autoWidth = TRUE))

if (length(num_display) > 0) {
  dt <- dt %>% formatRound(columns = num_display, digits = 1)
}

dt
```

Guide
=======================================================================

Row
-----------------------------------------------------------------------

### How This Dashboard Is Built

**Base.** This report is built fresh from `Book1.xlsx` every time it is
knitted — currently **`r comma(total_students)`** students after removing
duplicates and rows missing more than 60% of their answers. Personal
identifiers (name, ID, timestamp, email, phone) are dropped before any
analysis runs.

**Pillar construction.** Every pillar is the mean of its member survey
items on their original 1–5 scale, then multiplied by 20 so pillars run
from 20 to 100 and are directly comparable. Higher is always better. A
pillar only appears on this dashboard if at least 50 students answered
enough of its items to compute a score.

**Overall index.** The unweighted mean of every pillar shown on the
Executive Summary page — no pillar counts more than another.

**Net agreement** (Statement Diagnostics page). Share of students who
answered Agree or Strongly agree, minus the share who answered Disagree
or Strongly disagree, for that single statement. Neutral answers sit
outside the calculation, which is why the two shares don't need to add
to 100%.

**Importance** (Priority Matrix). The correlation between one statement
and the overall index. A high positive value means students who scored
that statement highly also tended to have a high overall index — it
describes an association, not a cause.

**Recommended action.**

- **Fix first** — below-average performance, above-average importance
- **Protect strength** — above-average performance and importance
- **Monitor** — below-average on both
- **Maintain** — above-average performance, below-average importance

**Risk segmentation.** Students are split into up to five bands —
Critical, At risk, Stable, Thriving, Excelling — using the 20th, 40th,
60th and 80th percentiles of their own overall index. Because the cut
points come from the data itself, segment sizes stay meaningful whatever
shape the distribution takes.

----

### Limits Worth Stating

- **Not a probability sample.** Respondents self-selected into this
  survey, so results describe this sample, not every student everywhere.
- **Association, not cause.** Every relationship on the Priority Matrix
  and Relationships page is correlational.
- **Small institutions move on a single answer.** Read the student count
  (n) beside every institution score before treating a rank as final.
- **Self-reported throughout.** No answer here was verified against
  administrative records.

### Keeping It Live

Replace `Book1.xlsx` with an updated export (same file name, same
folder), then knit this file again — every chart and table recalculates
automatically. Nothing on this dashboard is a hardcoded number.
